---
title: "Past Use of Internet - Group12"
author: "Gunjan Kumbhani, Pranjal Prakash, Anushka Shah, Zarin Tasnim, Vi Velumani"
date: "2022-11-30"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The Internet's impact and importance in people's day to day lives
continues to grow and thus, analyzing people's demographic attributes
with respect to their internet use could give us insight on internet use
behavior.

Within this analysis, we aim to explore data from the pastUse.csv data
file which contains data pertaining to 23,178 people's demographics,
education levels, household types, and past internet use. We
specifically focused on the relationships between a certain demographic
and a certain internet use category: Demographics analyzed in this
project: province, gender, age, labour force, education level Internet
Use Category analyzed in this project: frequency of internet use, last
time of internet use, and the reasons for non-internet use

## Pre-processing / Cleaning

This below code snippet shows the pre-processing of the data such that
we changed the numerical values of rows and columns to non-numerical
values. This helps us contextualize and understand the data and
visualizations that follow.

```{r preprocessingDataframes}

#loading all the packages
library(dplyr)
library(tidyverse)
library(epiDisplay)
library(ggplot2)

#Reading and storing data into two different dataframes (one raw, one for pre-processing)
PU <- read.csv("pastUse.csv", header = T)
df <- read.csv("pastUse.csv", header = T)

#Getting rid of the ID column
df <- df[ , -1]


#pre-processing the df dataframe so that column names and row values are non-numerical

df <- rename(df, isProvince = PROVINCE)  #change the name of the column
#df <- mutate(df, totalEdu = G_CEDUC + G_CSTUD) 

colRecode1 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 10, "Newfoundland and labrador", df[, i])
    df[, i] <<- ifelse(df[, i] == 11, "Prince Edward Island", df[, i])
    df[, i] <<- ifelse(df[, i] == 12, "Nova Scotia", df[, i])
    df[, i] <<- ifelse(df[, i] == 13, "New Brunswick", df[, i])
    df[, i] <<- ifelse(df[, i] == 24, "Quebec", df[, i])
    df[, i] <<- ifelse(df[, i] == 35, "Ontario", df[, i])
    df[, i] <<- ifelse(df[, i] == 46, "Manitoba", df[, i])
    df[, i] <<- ifelse(df[, i] == 47, "Saskatchewan", df[, i])
    df[, i] <<- ifelse(df[, i] == 48, "Alberta", df[, i])
    df[, i] <<- ifelse(df[, i] == 59, "British Columbia", df[, i])
  }
}

colRecode1 (1,1) 

df <- rename(df, isRegion = REGION)  #change the name of the column
#df <- mutate(df, totalEdu = G_CEDUC + G_CSTUD) 

colRecode2 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Atlantic Region", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "Quebec", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "Ontario", df[, i])
    df[, i] <<- ifelse(df[, i] == 4, "Manitoba / Saskatchewan", df[, i])
    df[, i] <<- ifelse(df[, i] == 5, "Alberta", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "British Columbia", df[, i])
  }
}

colRecode2 (2,2) 

df <- rename(df, isCommunity = G_URBRUR)  #change the name of the column
#df <- mutate(df, totalEdu = G_CEDUC + G_CSTUD) 

colRecode3 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 01, "Montreal", df[, i])
    df[, i] <<- ifelse(df[, i] == 02, "Toronto", df[, i])
    df[, i] <<- ifelse(df[, i] == 03, "Vancouver", df[, i])
    df[, i] <<- ifelse(df[, i] == 04, "Other urbcan excluding Prince Edward Island", df[, i])
    df[, i] <<- ifelse(df[, i] == 05, "Rural excluding Prince Edward Island", df[, i])
    df[, i] <<- ifelse(df[, i] == 06, "Prince Edward Island", df[, i])
  }
}

colRecode3 (3,3) 

df <- rename(df, isAgeGroups = GCAGEGR6)  #change the name of the column
#df <- mutate(df, totalEdu = G_CEDUC + G_CSTUD) 

colRecode4 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 01, "16 to 24", df[, i])
    df[, i] <<- ifelse(df[, i] == 02, "25 to 34", df[, i])
    df[, i] <<- ifelse(df[, i] == 03, "35 to 44", df[, i])
    df[, i] <<- ifelse(df[, i] == 04, "45 to 54", df[, i])
    df[, i] <<- ifelse(df[, i] == 05, "55 to 64", df[, i])
    df[, i] <<- ifelse(df[, i] == 06, "64 and older", df[, i])
  }
}

colRecode4 (4,4) 

df <- rename(df, isGender = CSEX)  #change the name of the column
#df <- mutate(df, totalEdu = G_CEDUC + G_CSTUD) 

colRecode5 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Male", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "Female", df[, i])
  }
}

colRecode5 (5,5) 

df <- rename(df,  isRespondentHighestEducationLevel = G_CEDUC)  #change the name of the column

colRecode6 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "High school or less", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "College or some post-secondary", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "University certificate or degree", df[, i])
  }
}

colRecode6 (6,6) 

df <- rename(df,  isStudentRespondent = G_CSTUD)  #change the name of the column

colRecode7 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
  }
}

colRecode7 (7,7) 

df <- rename(df,  isHouseholdHighestEduc = G_HEDUC)  #change the name of the column

df <- rename(df,  isStudentinHH = G_HSTUD)  #change the name of the column

colRecode12 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
  }
}

colRecode12 (12,12) 

df <- rename(df,  isEVPersonalUse = EV_Q01)  #change the name of the column

colRecode13 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
  }
}

colRecode13 (13,13) 

df <- rename(df,  isLabourForce = G_CLFSST)  #change the name of the column

colRecode8 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Employed", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "Unemployed", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "Not in the labour force", df[, i])
  }
}

colRecode8 (8,8) 

df <- rename(df,  isHHType = GFAMTYPE)  #change the name of the column

colRecode9 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Single family HH with unmarried children under 16", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "Single family HH without unmarried children under 16", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "One person HH", df[, i])
    df[, i] <<- ifelse(df[, i] == 4, "Multi family HH", df[, i])
  }
}

colRecode9 (9,9) 

df <- rename(df,  isNoofpersoninHH = G_HHSIZE)  #change the name of the column

colRecode11 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 4, "4 or more", df[, i])
  }
}

colRecode11 (11,11) 

df <- rename(df,  isInternetuseYears = EV_Q02)  #change the name of the column

colRecode14 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Less than a year", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "1 to 2 years", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "2 to 5 years", df[, i])
    df[, i] <<- ifelse(df[, i] == 4, "4 or more years", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid skip", df[, i])
    df[, i] <<- ifelse(df[, i] == 7, "Dont know", df[, i])
    df[, i] <<- ifelse(df[, i] == 8, "Refusal", df[, i])
  }
}

colRecode14 (14,14) 

df <- rename(df,  isPUPersonalUse = PU_Q01)  #change the name of the column

colRecode15 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid Skip", df[, i])
  }
}

colRecode15 (15,15)

df <- rename(df,  isPULastTimeUsed = PU_Q02)  #change the name of the column

colRecode16 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "1 to 2 Years", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "2 to 5 years", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "5 or more years", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid Skip", df[, i])
    df[, i] <<- ifelse(df[, i] == 7, "Dont Know", df[, i])
  }
}

colRecode16 (16,16)

df <- rename(df,  isPUHowOften = PU_Q03)  #change the name of the column

colRecode17 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "Atleast once a day", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "Atleast once a week", df[, i])
    df[, i] <<- ifelse(df[, i] == 3, "Atleast once a month", df[, i])
    df[, i] <<- ifelse(df[, i] == 4, "Less than once a month", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid Skip", df[, i])
    df[, i] <<- ifelse(df[, i] == 7, "Dont Know", df[, i])
  }
}

colRecode17 (17,17)

df <- rename(df,  isTooCostly = PU_Q06A)  #change the name of the column
df <- rename(df,  isNoNeed = PU_Q06E)  #change the name of the column
df <- rename(df,  isNoTime = PU_Q06J)  #change the name of the column
df <- rename(df,  isNoComputerAccess = PU_Q06K)  #change the name of the column

colRecode18 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid Skip", df[, i])
    df[, i] <<- ifelse(df[, i] == 7, "Dont Know", df[, i])
  }
}

colRecode18 (18,21)

df <- rename(df,  isOther = PU_G06)  #change the name of the column

colRecode22 <- function(startCol, endCol) {
  for (i in startCol:endCol) {
    df[, i] <<- ifelse(df[, i] == 1, "YES", df[, i])
    df[, i] <<- ifelse(df[, i] == 2, "NO", df[, i])
    df[, i] <<- ifelse(df[, i] == 6, "Valid Skip", df[, i])
    df[, i] <<- ifelse(df[, i] == 9, "Not Stated", df[, i])
  }
}

colRecode22 (22,22)
```

## How Long Have People Been Using the Internet?

To start off, getting an idea of the length of internet use among the
data helps us prove or disprove the claim that the internet plays a
major role in people's lives.

Through this distribution chart, we can see that 13,461 people out of
23,178 have been using the internet for 4 years or more. This means that
more than half of the respondents (58%) have been using the internet for
the longest amount of time, proving that the internet does play a major
role in the majority of people's lives.

```{r pressure}
#Showing how long people have been using the internet
tab1(df$isInternetuseYears, sort.group = "decreasing", cum.percent = TRUE, main = "Respondent Distribution of Years of Internet Use")

```

## Age and Gender Demographics for Province

In order to help us analyze internet use between age, gender, and
province, we must first take a look at the age and gender makeup of each
province.

From the "Age Group of Provinces" chart, we can that the most amount of
people live in Ontario while the least live in Prince Edward Island.
Moreover, we can see that people aged 45 years old and above makeup most
of the population of each province. From the "Gender Count by Province"
Chart, we can see that there are more females in each province than
there are males.

From the two charts combined, it is safe to say that for any given
province, the likelihood of encountering a female who is 45 years old
and above is more likely than encountering a male who is younger than 45
years old. This gives us a base to correlate gender, age, and province
based on internet use in the following sections.

```{r}
#Stacked bar chart showing age group of provinces
ageProv <- ggplot(data = df, aes(x = isProvince, fill =isAgeGroups))
ageProv <- ageProv + geom_bar()
ageProv <- ageProv + theme_minimal()
ageProv <- ageProv  + theme(axis.text.x = element_text(angle = 90,hjust = 0)) + labs(title="Age Group of Provinces")
ageProv

#Stacked bar chart showing gender count by province
freqprov_gen <- as.data.frame(table(df$isProvince, df$isGender))
freqprov_gen <- rename(freqprov_gen,  Province = Var1)
freqprov_gen <- rename(freqprov_gen,  Gender = Var2)
freqprov_gen <- rename(freqprov_gen,  Frequency_of_Respondents = Freq)

genProv <- ggplot(freqprov_gen, aes(x=Province, y=Frequency_of_Respondents, fill=Gender, label=Frequency_of_Respondents))+geom_bar(stat="identity")+geom_text(size=3, position=position_stack(vjust=0.5))+scale_fill_manual(values = c("Pink", "Lightblue"))+labs(title="Gender Count by Province")+theme(axis.text.x = element_text(angle = 90,hjust =0 ))
genProv
```

## Internet Use and Labour based on Province

Employment type can also influence internet use, so the following jitter
plots aim to investigate the correlation between them based on Province.

The "Labour Force By Province" plot shows that most people in each
province are employed. The "Internet Use By Province" plot shows that
most people in each province do use the internet. From this, we can
understand that people who are employed are more likely to use the
internet than people who are not employed.

```{r}
#Jitter Plot showing the labour force per province
laboProv <- ggplot(df, aes(y = isLabourForce, x = isProvince)) + geom_jitter() + theme(axis.text.x = element_text(angle = 90,hjust = 0)) + labs(title = "Labour Force By Province")
laboProv

#Jitter Plot showing internet use per province
intProv <- ggplot(df, aes(y = isPUPersonalUse, x = isProvince)) + geom_jitter() + theme(axis.text.x = element_text(angle = 90,hjust = 0)) + labs(title = "Internet Use By Province")
intProv
```

## Internet Use Behaviour Based on Gender

Investigating the relationship between internet use behaviour and gender
can give us insight as to how people are using the internet if they male
or female.

The "Frequency of Internet Use By Gender" chart shows that most people
use the internet atleast once a week. The distributions between male and
female are relatively uniform, although in the "atleast once a day"
category, more females tend to use the internet atleast once a day than
males.

The "Last Time Internet Use by Gender" chart shows that most people have
used the internet in the last 1-2 years, with the distribution of males
and females being relatively uniform.

From this, we can gather that gender does not play a major role in
determining frequency and recency of internet use.

```{r}
#Stacked Bar chart for showing frequency of internet use by Gender
trimmedHO <- filter(df, isPUHowOften!='Valid Skip' & isPUHowOften!='Dont Know')
freqHO_gen <- as.data.frame(table(trimmedHO$isPUHowOften, trimmedHO$isGender))
freqHO_gen <- rename(freqHO_gen,  Internet_Use = Var1)
freqHO_gen <- rename(freqHO_gen,  Gender = Var2)
freqHO_gen <- rename(freqHO_gen,  Frequency_of_Respondents = Freq)
howoften_GEN <- ggplot(freqHO_gen, aes(x=Internet_Use, y=Frequency_of_Respondents, fill=Gender, label=Frequency_of_Respondents))+geom_bar(stat="identity")+geom_text(size=5, position=position_stack(vjust=0.5))+scale_fill_manual(values = c("Pink", "Lightblue"))+theme(panel.background = element_rect(fill = "#887DA7", color = "black"))+labs(title="Frequency of Internet Use By Gender")
howoften_GEN

#Stacked Bar chart for showing last time internet use by Gender
trimmedLU <- filter(df, isPULastTimeUsed!='Valid Skip' & isPULastTimeUsed!='Dont Know')
freqLU_gen <- as.data.frame(table(trimmedLU$isPULastTimeUsed, trimmedLU$isGender))
freqLU_gen <- rename(freqLU_gen,  Last_Internet_Use = Var1)
freqLU_gen <- rename(freqLU_gen,  Gender = Var2)
freqLU_gen <- rename(freqLU_gen,  Frequency_of_Respondents = Freq)
lastuse_GEN <- ggplot(freqLU_gen, aes(x=Last_Internet_Use, y=Frequency_of_Respondents, fill=Gender, label=Frequency_of_Respondents))+geom_bar(stat="identity")+geom_text(size=5, position=position_stack(vjust=0.5))+scale_fill_manual(values = c("Pink", "Lightblue"))+theme(panel.background = element_rect(fill = "#7BA66E", color = "black"))+labs(title="Last Time Internet Use By Gender")
lastuse_GEN
```

## Reasons for Non-Internet Use based on Education

The level of education may determine why a person might not use the
internet. For example, in the "Too Costly?" Chart, we can see that the
most educated people (University degree) were the least to not use the
internet for expense purposes. From this, we can infer that this is due
to them being more financially stable since they have the most
education.

Similarly, in the "No Need" chart, we can see that the least educated
people (high-school or less) have the least need to use the internet,
which checks out since they might not use the internet as much as more
educated people might.

```{r}
#The following stacked bar charts show each reason for non-internet use based
#on the respondent's highest education level

#Too Costly
trimmedcostly <- filter(df, isTooCostly!='Valid Skip' & isTooCostly!='Dont Know')
freqcostly_highed <- as.data.frame(table(trimmedcostly$isTooCostly, trimmedcostly$isRespondentHighestEducationLevel))
freqcostly_highed <- rename(freqcostly_highed,  Too_Costly = Var1)
freqcostly_highed <- rename(freqcostly_highed,  Respondent_Highest_Education = Var2)
freqcostly_highed <- rename(freqcostly_highed,  Frequency_of_Respondents = Freq)
costly_highed <- ggplot(freqcostly_highed, aes(x=Respondent_Highest_Education, y=Frequency_of_Respondents, fill=Too_Costly, label = Frequency_of_Respondents)) + geom_bar(stat="identity") + geom_col() +
  geom_text(size=5, position=position_stack(vjust=0.5))+
  scale_x_discrete(guide = guide_axis(n.dodge=4))+
  scale_fill_manual(values = c("coral1", "darkseagreen"))+
  theme(panel.background = element_rect(fill = "#FCFBCC", color = "black"))+
  labs(title="Reason: Too Costly?") 
costly_highed


#No Need
trimmednoneed <- filter(df, isNoNeed!='Valid Skip' & isNoNeed!='Dont Know')
freqneed_highed <- as.data.frame(table(trimmednoneed$isNoNeed, trimmednoneed$isRespondentHighestEducationLevel))
freqneed_highed <- rename(freqneed_highed,  No_Need = Var1)
freqneed_highed <- rename(freqneed_highed,  Respondent_Highest_Education = Var2)
freqneed_highed <- rename(freqneed_highed,  Frequency_of_Respondents = Freq)
need_highed <- ggplot(freqneed_highed, aes(x=Respondent_Highest_Education, y=Frequency_of_Respondents, fill=No_Need, label = Frequency_of_Respondents)) + geom_bar(stat="identity") + geom_col() +
  geom_text(size=5, position=position_stack(vjust=0.5))+
  scale_x_discrete(guide = guide_axis(n.dodge=4))+
  scale_fill_manual(values = c("coral1", "darkseagreen"))+
  theme(panel.background = element_rect(fill = "#FCFBCC", color = "black"))+
  labs(title="Reason: No Need?") 
need_highed


#No Time
trimmednotime <- filter(df, isNoTime!='Valid Skip' & isNoTime!='Dont Know')
freqtime_highed <- as.data.frame(table(trimmednotime$isNoTime, trimmednotime$isRespondentHighestEducationLevel))
freqtime_highed <- rename(freqtime_highed,  No_Time = Var1)
freqtime_highed <- rename(freqtime_highed,  Respondent_Highest_Education = Var2)
freqtime_highed <- rename(freqtime_highed,  Frequency_of_Respondents = Freq)
time_highed <- ggplot(freqtime_highed, aes(x=Respondent_Highest_Education, y=Frequency_of_Respondents, fill=No_Time, label = Frequency_of_Respondents)) + geom_bar(stat="identity") + geom_col() +
  geom_text(size=5, position=position_stack(vjust=0.5))+
  scale_x_discrete(guide = guide_axis(n.dodge=4))+
  scale_fill_manual(values = c("coral1", "darkseagreen"))+
  theme(panel.background = element_rect(fill = "#FCFBCC", color = "black"))+
  labs(title="Reason: No Time?") 
time_highed

#No Computer
trimmednocomputer <- filter(df, isNoComputerAccess!='Valid Skip' & isNoComputerAccess!='Dont Know')
freqcomp_highed <- as.data.frame(table(trimmednocomputer$isNoComputerAccess, trimmednocomputer$isRespondentHighestEducationLevel))
freqcomp_highed <- rename(freqcomp_highed,  No_Computer = Var1)
freqcomp_highed <- rename(freqcomp_highed,  Respondent_Highest_Education = Var2)
freqcomp_highed <- rename(freqcomp_highed,  Frequency_of_Respondents = Freq)
comp_highed <- ggplot(freqcomp_highed, aes(x=Respondent_Highest_Education, y=Frequency_of_Respondents, fill=No_Computer, label = Frequency_of_Respondents)) + geom_bar(stat="identity") + geom_col() +
  geom_text(size=5, position=position_stack(vjust=0.5))+
  scale_x_discrete(guide = guide_axis(n.dodge=4))+
  scale_fill_manual(values = c("coral1", "darkseagreen"))+
  theme(panel.background = element_rect(fill = "#FCFBCC", color = "black"))+
  labs(title="Reason: No Computer?") 
comp_highed

```



```{r}
cm <- read.csv("pastUse.csv", header = T)
View(cm)

res <- cor(cm)
round(res, 2)

library("Hmisc")

res2 <- rcorr(as.matrix(cm))
res2

# Extract the correlation coefficients
res2$r
# Extract p-values
res2$p

install.packages("corrplot")
library(corrplot)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)

```
